---
title: "Exploring Fitness"

output: 
  html_document:
    keep_md: true

---


```{r setoptions, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
options(scipen = 999)
```

## Introduction

In the following report, I run a quick exploratory analysis of the fitness data.
I will be using the Tidyverse package for plotting and manipulating the data.

``` {r load, results="hide", message=FALSE}
library(tidyverse)
t1<-read.csv("activity.csv")
```

## Part 1: Mean, Median, Histogram 

``` {r Part1}
sum(t1$steps, na.rm = T)
smry<-tapply(t1$steps, t1$date, sum, na.rm=TRUE)
mean(smry, na.rm=TRUE)
median(smry, na.rm=TRUE)
smrytbl<-as.data.frame(smry)
smrytbl %>% ggplot(aes(smry)) + geom_histogram(bins=20)
```

Excluding the NA values:

- the total number of steps is `r sum(t1$steps, na.rm = T)`. 
- the  mean number of steps per day is `r mean(smry, na.rm=TRUE)`. 
- the median number of steps per day is `r median(smry, na.rm=TRUE)`.

## Part 2: Time Series For Average Interval

``` {r Part2}
tsplot<-t1 %>% group_by(interval) %>% summarise(mean=mean(steps,na.rm = T)) 
tsplot %>% ggplot(aes(interval, mean))+geom_line()
tsplot$interval[which.max(tsplot$mean)]
```

The time series plot shows a clear spike after 800, and the busiest interval is 835, 
with around `r round(tsplot$mean[which.max(tsplot$mean)])` steps.

## Part 3: Replacing NA Values

To replace NA values, I merged the table containing average values per interval with the original table, and replaced all NA values with the corresponding interval average.

```{r Part3}
sum(is.na(t1$steps))
t2<-(merge(tsplot, t1))
t2<-t2[order(t2$date), ]
t2$steps[is.na(t2$steps)]<-t2$mean[is.na(t2$steps)]

smry2<-tapply(t2$steps, t2$date, sum)
mean(smry2)
median(smry2)
smrytbl2<-as.data.frame(smry2)
smrytbl2 %>% ggplot(aes(smry2)) + geom_histogram(bins=20)
```

There are `r sum(is.na(t1$steps))` NA values.

After replacing the NA values:

- the total number of steps is `r round(sum(t2$steps))`. 
- the  mean number of steps per day is about `r round(mean(smry2))`. 
- the median number of steps per day is `r (median(smry2))`.


## Part 4: Weekday vs Weekend

In this part, I create a factor variable for the weekdays and weekends, and then use ggplot2 to create panel plots, faceting on the factor variable.

```{r Part4}
t3<-t2
t3$date<-as.Date(t3$date)
t3$weekdays<-weekdays(t3$date)
t3$ends<-factor(t3$weekdays)
levels(t3$ends)<-c("Weekday", "Weekday", "Weekend", "Weekend", "Weekday", "Weekday", "Weekday")

t3plot<-t3 %>% group_by(interval, ends)%>%summarise(mean=mean(steps))
t3plot %>% ggplot(aes(interval, mean))+geom_line()+facet_wrap(~ends)

```
